\chapter{Bibliograf\'ia}\label{Bibliografia}
\begin{enumerate}
\item Agarwal, Apoorv; Xie, Boyi; Vovsha, Ilia; Rambow, Owen; Passonneau, Rebecca. Columbia University, 2011. Art\'iculo, LSM '11 ``Proceedings of the Workshop on Languages in Social Media'', P\'aginas 30-38, Association for Computational Linguistics Stroudsburg, PA, USA. ``Sentiment Analysis of Twitter Data''.
\item Alves Kaestner, Celso Antonio. Federal University of Technology, Curitiba, PR, Brasil, 2013. Tutorial, Revista de Inform\'atica Te\'orica e Aplicada - RITA, Volumen 20, N\'umero 3, P\'aginas 130-154,  Instituto de Inform\'atica da Universidade Federal do Rio Grande do Sul, Brasil. ``Support Vector Machines and Kernel Functions for Text Processing''. 
\item Baeza-Yates, Ricardo; Ribeiro-Neto, Berthier. ``Modern Information Retrieval''. USA: ACM Press, 1999.
\item Bal, Daniella; Bal, Malissa; Hogenboom, Alexander; Hogenboom, Frederik; Frasincar, Flavius. Erasmus University Rotterdam, DR, Rotterdam; van Bunningen, Arthur. Teezir BV, KL, Utrecht, 2011. Publicaci\'on, WISE 2011 Proceedings of the 12th International Conference on Web Information System Engineering, P\'aginas 129-142, Springer-Verlag, Berlin, Heidelberg. ``Sentiment Analysis with a Multilingual Pipeline''.
\item Bautin, Mikhail; Skiena, Steven. Stony Brook University, NY, USA; Vijayarenu, Lohit. Yahoo! Inc., CA, USA, 2008. Publicaci\'on, ICWSM 2008 2nd International Conference on Weblogs and Social Media, P\'aginas 19-26, AAAI Press. ``International Sentiment Analysis for News and Blogs''.
\item Boiy, Erik; Moens, Marie-Francine. Katholieke Universiteit Leuven, Belgium, 2008. Publicaci\'on, Information Retrieval. Volumen 12 N\'umero 5, P\'aginas 526-558, Kluwer Academic Publishers Hingham, MA, USA (2009). ``A Machine Learning Approach to Sentiment Analysis in Multilingual Web Texts''.
\item Brody, Samuel; Diakopoulos, Nicholas. Rutgers University, 2011. Publicaci\'on, EMNLP '11 Proceedings of the Conference on Empirical Methods in Natural Language Processing, P\'aginas 562-570, Association for Computational Linguistics Stroudsburg, PA, USA.
\newline ``Cooooooooooooooollllllllllllll!!!!!!!!!!!!!!: using word lengthening to detect sentiment in microblogs''.
\item Burges, Christopher J.C. Bell Laboratories, Lucent Technologies, Lucent Technologies, 1998. Art\'iculo, Data Mining and Knowledge Discovery Volumen 2 N\'umero 2 (1998) P\'aginas 121-167. Kluwer Academic Publishers Hingham, MA, USA. ``A Tutorial on Support Vector Machines for Pattern Recognition''.
\item Chang, Chih-Chung; Lin, Chih-Jen. Department of Computer Science, National Taiwan University, Taipei, Taiwan, 2001. Art\'iculo, ACM Transactions on Intelligent Systems and Technology Volumen 2 N\'umero 3 (2011) Art\'iculo 27, ACM New York, NY, USA. ``LIBSVM: A library for support vector machines''. 
\item Ding, Xiaowen; Liu, Bing; Yu, Philip S. University of Illinois at Chicago, Chicago, IL, 2008. Publicaci\'on, WSDM '08 Proceedings of the 2008 International Conference on Web Search and Data Mining, P\'aginas 231-240, ACM New York, NY, USA. ``A Holistic Lexicon-Based Approach to Opinion Mining''.
\item Fayyad, Usama; Piatetsky-Shapiro; Smyth, Padhraic. Publicaci\'on, Advances in knowledge discovery and data mining, P'aginas 1-34, American Association for Artificial Intelligence Menlo Park, CA, USA, 1996. ``From Data Mining to Knowledge Discovery''.
\item Friedman, Jerome; Hastie, Trevor; Tibshirani, Robert. Stanford University, 1998. Art\'iculo, The Annals of Statistics Volumen 28 N\'umero 2 (2000), P\'aginas 337-407. ``Additive Logistic Regression: a Statistical View of Boosting''.
\item Gim\'enez-Lugo, Gustavo. Universidade de S\~ao Paulo. ``Um Modelo de Sistemas Multiagentes para Partilha de Conhecimento Utilizando Redes Sociais Comunit\'arias''. Tesis de Doctorado en Ingenier\'ia, 2004.
\item Go, Alec; Bhayani, Richa; Huang, Lei. Reporte t\'ecnico, Stanford University, 2009. ``Twitter Sentiment Classification using Distant Supervision''.
\item Hu, Minqing; Liu, Bing. University of Illinois at Chicago, Chicago, IL, 2004. Publicaci\'on, KDD '04 Proceedings of the tenth ACM SIGKDD international conference on Knowledge discovery and data mining, P\'aginas 168-177, ACM New York, NY, USA. ``Mining and summarizing customer reviews''.
\item Joachims, Torsten. Universit\"at Dortmund, Germany, 1998. Publicaci\'on, ECML '98 Proceedings of the 10th European Conference on Machine Learning, P\'aginas 137-142, Springer-Verlag London, UK. ``Text Categorization with Suport Vector Machines: Learning with Many Relevant Features''.
\item Kennedy, Alistair; Inkpen, Diana. University of Ottawa, Ottawa, Canada, 2006. Art\'iculo, Computational Intelligence, Volumen 22, N\grado\ 2, Mayo 2006, P\'aginas 110-125. ``Sentiment Classification of Movie Reviews Using Contextual Valence Shifters''.
\item Kim, Soo-Min; Hovy, Eduard. University of Southern California, Marina del Rey, CA, 2004. Publicaci\'on, COLING '04 Proceedings of the 20th international conference on Computational Linguistics, Art\'iculo N\grado \ 1367, Association for Computational Linguistics Stroudsburg, PA, USA. ``Determining the sentiment of opinions''.
\item Landwehr, Niels. University of Freiburg, Germany. Hall, Mark; Frank, Eibe. University of Waikato, New Zealand, 2005. Art\'iculo, Machine Learning Volumen 59 N\'umero 1-2 (2005), P\'aginas 161-205, Kluwer Academic Publishers, Hingham, MA, USA. ``Logistic Model Trees''.
\item Le Cessie, S; Van Houwelingen J.C. University of Leiden, The Netherlands, 1990. Art\'iculo, Applied Statistics Volumen 41 N\'umero 1 (1992), P\'aginas 191-201. ``Ridge Estimators in Logistic Regression''.
\item Lee, Huey Yee; Renganathan, Hemnaath. Jalan Universiti, Ehsan, Malaysia, 2011. Publicaci\'on, SAAIP 2011 Proceedings of the Workshop on Sentiment Analysis where AI meets Psychology, P\'aginas 89-93, Asian Federation of Natural Language Processing. ``Chinese Sentiment Analysis Using Maximum Entropy''.
\item Liu, Bing. Morgan \& Claypool Publishers, Mayo 2012. ``Sentiment Analysis and Opinion Mining''.
\item Lustig, Wolf. Mainz, 1996. Ensayo. ``Mba'\'eichapa oiko la guarani? Guaran\'i y jopara en el Paraguay''.
\item Meli\'a Lliteres, Bartomeu. Colecciones Mapfre, Madrid, 1992. ``La lengua guaran\'i del Paraguay: Historia, sociedad y literatura (Lenguas y Literaturas Ind\'igenas)''.
\item Moreno-Ortiz, Antonio; P\'erez-Hern\'andez, Chantal. Universidad de M\'alaga, 2013. Art\'iculo, Procesamiento del Lenguaje Natural, N\grado\ 50, P\'aginas 93-100, Sociedad Espa\~nola para el Procesamiento del Lenguaje Natural. ``Lexicon-Based Sentiment Analysis of Twitter Messages in Spanish''.
\item Nigam, Kamal. Carnegie Mellon University, Pittsburgh, PA; Lafferty, John. Carnegie Mellon University Pittsburgh, PA; McCallum, Andrew. Just Research, Pittsburgh, PA, 1999. Publicaci\'on, IJCAI-99 Workshop on Machine Learning for Information Filtering, P\'aginas 61-67. ``Using maximum entropy for text classification''.
\item Pak, Alexander; Paroubek, Patrick. Universit\'e de Paris-Sud, 2010. Publicaci\'on. LREC '10 Proceedings of the Seventh International Conference on Language Resources and Evaluation. ``Twitter as a Corpus for Sentiment Analysis and Opinion Mining''.
\item Pang, Bo. Cornell University, Ithaca, NY; Lee, Lillian. Cornell University, Ithaca, NY; Vaithyanathan, Shivakumar. IBM Almaden Research Center, San Jose, CA, 2002. Publicaci\'on, EMNLP '02 Proceedings of the ACL-02 conference on Empirical methods in natural language processing - Volumen 10, P\'aginas 79-86, Association for Computational Linguistics Stroudsburg, PA, USA. ``Thumbs up?: sentiment classification using machine learning techniques''.
\item Pang, Bo; Lee, Lilian. Cornell University, Ithaca, NY, 2004. Publicaci\'on, ACL '04 Proceedings of the 42nd Annual Meeting on Association for Computational Linguistics, Art\'iculo N\grado \ 271, Association for Computational Linguistics Stroudsburg, PA, USA. ``A sentimental education: sentiment analysis using subjectivity summarization based on minimum cuts''.
\item Prasad, Suhaas. Reporte t\'ecnico, Stanford University, 2010. ``Micro-blogging Sentiment Analysis Using Bayesian Classification Methods''.
\item Saralegi Urizar, Xabier; San Vicente Roncal, I\~naki. Elhuyar Fundazioa, Espa\~na, 2012. Art\'iculo, TASS 2012 Workshop on Sentiment Analysis at SEPLN. ``TASS: Detecting Sentiments in Spanish Tweets''.
\item Taboada, Maite. Simon Fraser University; Brooke, Julian. University of Toronto; Tofiloski, Milan. Simon Fraser University; Voll, Kimberly. University of British Columbia; Stede, Manfred. University of Postdam, 2011. Art\'iculo, Journal Computational Linguistics Volume 37 Issue 2, Junio 2011, P\'aginas 267-307, MIT Press Cambridge, MA, USA. ``Lexicon-based methods for sentiment analysis''.
\item Turney, Peter D. National Research Council of Canada, Ottawa, Ontario, Canada, 2002. Publicaci\'on, ACL '02 Proceedings of the 40th Annual Meeting on Association for Computational Linguistics, P\'aginas 417-424, Association for Computational Linguistics Stroudsburg, PA, USA. ``Thumbs up or thumbs down?: semantic orientation applied to unsupervised classification of reviews''.
\item Vilares, David; Alonso Miguel A.; G\'omez-Rodr\'iguez, Carlos. Universidade da Coru\~na, A Coru\~na, 2013. Art\'iculo, Procesamiento del Lenguaje Natural, N\grado\ 50, P\'aginas 13-20, Sociedad Espa\~nola para el Procesamiento del Lenguaje Natural. ``Clasificaci\'on de polaridad en textos con opiniones en espa\~nol mediante an\'alisis sint\'actico de dependencias''.
\item Wan, Xiaojun. Peking University, Beijing, China, 2009. Publicaci\'on, ACL '09 Proceedings of the Joint Conference of the 47th Annual Meeting of the ACL and the 4th International Conference on Natural Language Processing of the AFNLP, Volumen 1, P\'aginas 235-243, ACL, Stroudsburg, PA, USA. ``Co-training for cross-lingual sentiment classification''.
\item Yang, Yiming; Liu, Xin. School of Computer Science, Carnegie Mellon University, Pittsburgh, PA, 1999. Publicaci\'on, SIGIR '99 Proceedings of the 22nd annual international ACM SIGIR conference on Research and development in information retrieval, P\'aginas 42-49, ACM New York, NY, USA. ``
A re-examination of text categorization methods''.
\item Zhang, Lei; Ghosh, Riddhiman; Dekhil, Mohamed; Hsu, Meichun; Liu, Bing. HP Laboratories, 2011. Reporte T\'ecnico, HPL-2011-89. ``Combining Lexicon-based and Learning-based Methods for Twitter Sentiment Analysis''.

\end{enumerate}